---
title: "Stockholm Housing Prices"
author: "Selja"
date: "1/9/2022"
output: html_document
---

## Setting up my enviroment
```{r loading packages}

library(sf)
library(tidyverse)
library(jsonlite)
library("dplyr") 
library("tidyr")
```
## Accesing the GEO location data for Sweden
```{r Accesing to the GEO location data for Sweden}

tmp <- tempfile()
download.file("http://api.thenmap.net/v2/se-7/geo/2020-06-06", destfile = tmp)

mun_name <- fromJSON("http://api.thenmap.net/v2/se-7/data/2020-06-06?language=sv&data_props=name|shapeid|is_in") %>% 
  unnest(is_in) %>% 
  rename(county = is_in)

```
## Joining the two datasets
```{r Joining the 2 datasets}
mun <- read_sf(tmp) %>% 
  left_join(mun_name, by = c("id" = "shapeid")) 

```
## Filtering the dataset to show only Stockholm area
```{r}
stockholm_mun <- mun %>% 
  filter(county == "Stockholms län") %>% 
  mutate(fill_data = rnorm(nrow(.))) 
```
#Plotting the dataset
```{r}
ggplot(stockholm_mun) +
  geom_sf(aes(fill = fill_data)) +
  scale_fill_viridis_c() +
  theme_void()
```
## Let's add the average housing prices. This data can be find at https://www.maklarstatistik.se/ and check the column names.

```{r}
library(readxl)

house_price <- read_excel("housing_stockholm_2021.xlsx")

colnames(house_price)
```
## Renaming some of the column names for merging and clarity
```{r}
names(house_price) [names(house_price) == "Områden"] <- "name"
names(house_price) [names(house_price) == "Kr/kvm"] <- "PricePerSqm"

View(house_price)
```
## Merging housing prices and geo data by name column

```{r}
total <- merge(stockholm_mun,house_price, by = "name")
```
## Plotting the data.

```{r}
ggplot(data = total) +
  geom_sf() +
  geom_sf(data = total, aes(fill = PricePerSqm)) +
  scale_fill_viridis_c(trans = "sqrt", alpha = .4) +
  theme_void()
```


## Let's break the Price per Sqm to intervals

```{r}
library(classInt)

breaks_qt <- classIntervals(c(min(total$PricePerSqm),0,total$PricePerSqm), n = 7, style = "quantile")

breaks_qt

total <- mutate(total, PricePerSqm_cat = cut(PricePerSqm, breaks_qt$brks)) 

total <- cbind(total, st_coordinates(st_centroid(total)))

```
## Plotting the data with new intervals
```{r}
ggplot(total) + 
  geom_sf(aes(fill=PricePerSqm_cat)) +
  geom_label(data = total, aes(X, Y, label = name), size = 3) +
  scale_fill_brewer(palette = "OrRd") 
```


## To avoid the overlapping, ggrepel library is applied
```{r}
library(ggrepel)
```
## Let's plot the result to see what is the outcome now.
```{r}
ggplot(total) + 
  geom_sf(aes(fill=PricePerSqm_cat)) +
  geom_text_repel(data = total, aes(X, Y, label = name))+
  scale_fill_brewer(palette = "OrRd") 
```


## Still quite hard to read, let's try without labels.

```{r}
ggplot(total) + 
  geom_sf(aes(fill = PricePerSqm_cat), show.legend = TRUE) + 
  geom_sf_label(aes(label = NA), label.padding = unit(1, "mm")) +
  scale_fill_brewer(palette = "OrRd")
```


## Let's try with another color scheme and some of the labels.

```{r}
ggplot(total) + 
  geom_sf(aes(fill=PricePerSqm_cat)) +
  geom_text_repel(data = total, aes(X, Y, label = name), nudge_y = c(0.04))+
  scale_fill_viridis_d()
```



### For future work, it would be handy to cut the amount of labels for readability
### Interval can be changed to enchance the readability
### Naming and other visual components should be added
### But this will do for now, such a fun Saturday project!
